\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{url}
\usepackage{listings}
\usepackage{graphicx} % Required for \resizebox
\usepackage{array} % Required for column width adjustments
\usepackage{tabularx}
\usepackage{balance}
\usepackage{colortbl} 
\usepackage{booktabs}
\usepackage{footnote}
\usepackage{multirow} 
 
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Automating Maritime Risk Data \\Collection and Identification Leveraging \\Large Language Models}

% \title{1.Classification Model for Global Incidents and Establishing Data Collection Pipeline for Disaster Research. \\2.Automating Maritime Risk Data Collection and Categorization\\
% 3 Automating Maritime Risk Data Collection and Categorization: Comparing Traditional and Advanced Models}
% Which title is better? Can select and edit to produce a good title.
%We prefer the second one "Automating Maritime Risk Data Collection and Categorization" among the 3 possible title

\author{Anonymous Submission}

%\iffalse
\author{\IEEEauthorblockN{Donghao HUANG\textsuperscript{1,2},
 Xiuju FU\textsuperscript{3*}, Xiaofeng YIN\textsuperscript{3}, Haibo PEN\textsuperscript{4*}, Zhaoxia WANG\textsuperscript{1*}
}

\IEEEauthorblockA{School of Computing and Information Systems, Singapore Management University, Singapore\textsuperscript{1}}
\IEEEauthorblockA{Research and Development, Mastercard, Singapore\textsuperscript{2}}
\IEEEauthorblockA{Institute of High Performance Computing, Agency for Science, Technology and Research, Singapore\textsuperscript{3}}
\IEEEauthorblockA{School of Electrical and Information Engineering, Tianjin University, China\textsuperscript{4}}
\IEEEauthorblockA{dh.huang.2023@smu.edu.sg; fuxj@ihpc.a-star.edu.sg*; yinxf@ihpc.a-star.edu.sg; penhaibo@tju.edu.cn*; zxwang@smu.edu.sg*}
%\thanks{\IEEEauthorrefmark{1}
%Corresponding Authors: Zhaoxia WANG (e-mail: zxwang@smu.edu.sg)
%}
}
%\fi
\maketitle

\begin{abstract}
%This paper presents an innovative method for automating the collection and categorization of global maritime incident news, addressing limitations of traditional manual methods. We evaluate various classification models, including conventional machine learning approaches and advanced Large Language Models (LLMs) like GPT-4o and Llama-3.1. Our results demonstrate that LLMs, particularly GPT-4o, consistently outperform traditional models, significantly improving both efficiency and accuracy of maritime risk data collection. We introduce a new evaluation metric, the "Ratio of Valid Categories," to assess model reliability. Our findings contribute to the expanding literature on LLM applications in risk management, demonstrating their potential to transform data collection practices.
%zx-edit to focus on the contributions of this research

Maritime risk research is crucial yet challenging for improving safety, efficiency, and sustainability in maritime operations. This paper presents an innovative method for automating the collection and identification of risk data related to global maritime risks from news sources, addressing the limitations of traditional manual methods.
To evaluate the proposed method, different learning-based models, including conventional machine learning approaches and advanced Large Language Models (LLMs) such as GPT-4 and LLaMA-3.1, are comprehensively studied for comparison. 
In addition, not only do we use popular evaluation metrics to assess the proposed method, but we also introduce a new evaluation metric, called the "Ratio of Valid Categories ($RVC$)," to evaluate model reliability. The merits of the proposed method are demonstrated across different evaluation metrics.
The research results show that the proposed LLM-based methods, particularly the GPT-4-based method, consistently outperform traditional models, significantly improving both the efficiency and accuracy of maritime risk data collection and identification. Our findings contribute to the expanding literature on LLM applications in risk management, demonstrating their potential to transform data collection and identification practices.

\end{abstract}

\begin{IEEEkeywords}
%Maritime Risk Assessment, Automated Data Collection, Incident Categorization, Large Language Models, Few-shot Learning, Traditional Machine Learning, GPT-4o, Llama-3.1, Text Classification, Ratio of Valid Categories 
%zx edit to remove the keywords, which do not appear  in the abstract
Maritime Risk, Automated Data Collection, Risk Identification, Large Language Models, Traditional Machine Learning, GPT-4o, Llama-3.1, Ratio of Valid Categories ($RVC$) 
\end{IEEEkeywords}

% \begin{IEEEkeywords}
% Data Collection Pipeline, Maritime Risk Research, Incident Categorization, Large Language Models, Support Vector Machine, Naive Bayes, Random Forest, K-nearest Neighbors, Text Classification
% \end{IEEEkeywords}

\input{intro}

\input{related_work}

\input{dataset}

\input{method}

\input{results}

\input{conclusion}

\section*{Acknowledgment}

The authors express their sincere appreciation to the students from Singapore Management University (SMU) for their enthusiastic interest and invaluable contributions to this paper: LU Quanfang, WEI Yifan, YANG Xinyue, and LIAO Jiaxiong. Their dedication has significantly enriched our research. This work was supported by the Natural Science Foundation
of Tianjin under Grant 19JCQNJC06000.

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
