\section{Introduction}
\subsection{Background and Motivation}
        The turtle soup game is a mystery puzzle made famous by Paul Sloane's book Lateral Thinking Puzzles and enjoyed by youths all around the world. It is named after a classic problem in the genre: a man ordered turtle soup and committed suicide afterwards. Players have to deduce the rationale behind the suicide by asking questions which can only be answered with yes or no. These problems promote parallel thinking: to break the frame of thought from a new direction or a different perspective on a phenomenon that does not make sense or is incomprehensible to common sense at first glance.  They represent the logical reasoning task in our paper.

    Logical Reasoning with Large Language Models (LLMs) has advanced significantly, but it remains a challenging and evolving area. LLMs like GPT-4 have stellar pattern recognition capabilities and are able to make inferences from large-scale text data. LLMs can handle basic formal logic tasks, such as propositional logic or syllogisms, and they can respond reasonably well to informal reasoning challenges like common-sense questions. LLMs have developed to understand context and leverage the nuanced relationships between sentences and concepts to make more accurate logical decisions. 
    
    However, LLMS still often struggle with deductive reasoning, where consistency is critical across multiple steps of reasoning. While they can generate logical conclusions in isolation, maintaining a chain of reasoning over several steps or sentences remains difficult. When dealing with multi-step complex reasoning, LLMs may make errors which compound over long logical chains as intermediate steps build upon one another.  LLMs also face limitations in tasks where reasoning requires abstract thought or the manipulation of symbolic or formal rules.
    
    Thanks to the Artificial Intelligence (AI) boom, there are more LLMs that have risen up to compete with traditional LLM powerhouses such as GPT by OpenAI and Llama by MetaAI. Open source LLMs like Qwen and Mistral have proven themselves to be worthy adversaries to their paid closed-source competitors. Alibaba Cloud recently released Qwen2, a series of successor models to the widely appreciated Qwen. Qwen2 boasts outstanding performances in both the large and small models. Qwen2 72B model outperforms the newly released Llama3 70B model and the mixture of exports Mixtral 8x22B model in benchmarks like MMLU, MMLU-Pro, TheoremQA, HumanEval, GSM8k and many more\cite{https://qwenlm.github.io/blog/qwen2/}. Meanwhile, Qwen2 7B Instruct Model surpasses the newly introduced State-Of-The-Art (SOTA) models like the Llama3 8B Model and the GLM4 9B Model despite being the smallest model of the three. All Qwen2 models are equipped with Grouped Query Attention, which greatly speeds up the process of attention and decreases the amount of memory required to store the intermediate results during inference.
    
    Qwen2 also represents the subset of LLMs developed in China. Their capabilities are of high interest to academia as the whole AI ecosystem stands to benefit from a more competitive and diverse global AI landscape, with reduced dominance by Western, especially US based, tech companies. Chinese LLMs are often optimised for tasks that require fluency in both Chinese and English (and sometimes other languages). This signifies exciting potential for cross-border trade, international diplomacy, and other areas where bilingual capabilities are valued. These models are more attuned to Chinese and Asian cultural, social, and historical contexts, which help address biases that might exist in models developed in the West.  

    We are interested in evaluating the newest generation of open source LLMs such as Qwen2 and Llama3.1 on their capabilities in answering complex and abstract logical reasoning tasks. We finetune LLMs on a variation of turtle soup Question Answering (QA) dataset in Chinese, where LLMs act as the game master and answer questions.
\subsection{Contributions}
Our main contributions are:
\begin{itemize}
    \item TODO
\end{itemize}
