\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

I would like to express my deepest gratitude to my supervisor, Associate Professor Zhaoxia WANG, for her invaluable guidance, continuous support, and patience throughout this research journey. Her expertise in artificial intelligence and natural language processing has been instrumental in shaping this dissertation.

I am also grateful to Professor Chong Wah NGO for his co-supervision and insightful feedback that helped refine the direction of this research. Special thanks to Associate Professor Qianru SUN for serving on my dissertation committee and providing constructive suggestions. I would also like to thank Professor Chunyan MIAO of Nanyang Technological University for her time and expertise as the external reviewer of this dissertation.

This research was supported in part by the School of Computing and Information Systems at Singapore Management University, which provided the academic environment and resources necessary for this work. I also gratefully acknowledge the support of the Singapore Digital Scholarship (Postgraduate), awarded to me in 2025 by the Infocomm Media Development Authority (IMDA).\footnote{\url{https://www.imda.gov.sg/how-we-can-help/singapore-digital-scholarship/postgraduate}}

I extend my appreciation to Mastercard for their support of this research. The collaboration with colleagues in Research and Development on retrieval-augmented generation (RAG), multi-agent systems, and edge deployment projects directly contributed to several publications in this dissertation. I am also grateful for access to computational resources, including GPU infrastructure for model training and evaluation, as well as the flexibility to present this research at international conferences.

I also thank my co-authors---Thanh-Son NGUYEN, Fiona LIAUSVIA, Joon Kiat CHUA, Samuel SAMUEL, James TAN Yong Hong, Gauri MALWE, Ritik VARSHNEY, Alvin TSE, Shila CHEW, Anna DUTKIEWICZ, Aparna CHEKURU, Erin DEUTSCHMAN, Zhenda HU, Nicole TEO, Prasan PAI, and others---for their contributions to the publications underlying this research.

Finally, I thank my family and friends for their unwavering support and encouragement throughout this doctoral journey.